Procedural generation of levels for the angry birds videogame using evolutionary computation\\

### 2019-05-16 IEEE_ConferenceOnGames/new_article.tex
few changes

\begin{abstract}
In this paper, we propose an evolutionary algorithm that follows an
open-ended evolution approach to generate levels for the Angry Birds video
game. The levels themselves are composed of a set of birds that the player
can use, a certain amount of pigs that they must destroy, and a given number
of pieces that conform structures that may or may not protect the pigs from
birds. The current goal is the generation of diverse structures playable in
the game, having the additional characteristics of being fun and enjoyable.
We propose a multi-layered search, first constructing composite structures
from basic blocks, to then build more complex structures building from these
composites, following an open-ended evolution approach in which the
evolution of structures is not guided towards a single objective but is
rather free to evolve and generate novelty or diversity. The fitness
function we use to evaluate the proposed levels considers how complex they
have become and how different they are from the rest of the group. The
experiments conducted show that an evolutionary approach allows the
generated levels to be novel and interesting to play.
% Another revision to English would be good.
% Done - Mario
\end{abstract}

@@

\section{State of the art}
\label{SoA}
In this Section, we present PCG works relevant to the level generation for
the Angry Birds video game. Renz et al. \cite{Renz, Renz2015TheAB} organized
an AI competition to develop intelligent agents that could play new  Angry
Birds levels better than the best human players. The competition was then
collocated in other AI conferences. To better train and test agents, more
novel and some times, challenging levels must be generated. Stephenson et
al. proposed a related competition this time with the goal of developing a
level generator for the game. Researchers have proposed several generators
in the following years of competition. M. Katina et al.  \cite{Kaidan2015}
propose a model in which the algorithm controls the generation of content by
using the skill of a player, while she is actively playing the game. The
moment the player finishes a certain level, the algorithm obtains the score
and generates the next level to play. Y. Jiang et al. \cite{Jiang2017}
proposed the use of predefined letter patterns and combined them to create a
small pool of words and letters used to create levels with text on them.
Users would play the game, and after that,   asked them if they liked the
layout of the Levels.  Several researchers follow a search-based approach.
Calle et al. \cite{lauracalle} propose a free-search method to generate
levels.  For this, lists of blocks are placed in the game by the algorithm,
and those maintaining their positions against gravity are selected to build
structures. The resulting structures continue evolving into more complex
levels. That work is the basis of our current proposal, and since we are
using a modified version of the Angry Birds game, that can obtain data of a
level after a simulation.
    %%%
    % TO DO: Laura´s work here 
    %%%

    \section{Proposed Method}
    \label{prop_method}  
    % Is this explanation undertandable ? JSH
    In this Section, we describe the approach used in this work, explaining
first, the restrictions of the game, then giving a quick overview of how we
create new composites to be used by the algorithm, then we define how candidate
solutions are to be represented by the algorithm. We then describe our method to
evaluate candidate solutions. Finally, we present details of the evolutionary
algorithm.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[width=80mm]{Images/list_pieces.png}}
\caption{List of blocks in the angry birds game.}
\label{piece_list}
\end{figure}
    % Explain a bit about the competition, from the introduction
The Angry Birds level generation track is a competition where the objective
is to generate levels that are \textit{fun} and \textit{enjoyable}; they also have to be challenging but not impossible. The rules for the
competition specify that a configuration file must be used to generate the
levels of the game in a limited amount of time. The file contains several
restrictions; namely:\begin{itemize}
\item   The number of \textit{different}
levels that must be generated. 
\item The materials that are \textit{not allowed}.

@@

    Structures are represented in a way similar to Togelius et al.
    \cite{togelius2016Representationsforsearch-basedmethods}. Each block has
    an ID assigned to it. A structure is created from a list of IDs, every one of them referring to
    composite blocks; the first 11 are basic composites, while generated composites 
    are assigned a higher number.  This way and individual can be
    represented as shown on figure \ref{old_chrom} where each chromosome
    represents a certain item in the list of blocks, the numbers for these
    elements are assigned in a first come first served basis; they can be infinitely incremented. The same blocks can be repeated in a structure.
    % I don't understand this too well, maybe rephrasing would clarify - JJ

@@

    These structures represent individuals in the evolutionary algorithm. Once
    an individual is created it is ready to be evaluated in a simulation on the
    game environment. %In Laura's work, we tried to avoid this because it took a long time. Is there any procedure to check constraints and not evaluate them? - JJ
    Using a similar approach as the composite generator, all
    individuals have a maximum of ten seconds to maintain balance or reach
    absolute stability. If that happens  it means one of two
    things, first the generated structure is completely stable or second, the
    tower fell to either side and the pieces were destroyed or stopped moving by
    gravitational pull, when either of this happens by more than two seconds the
    simulation ends regardless of time remaining and the next individual is
    simulated.

    The evaluation process for this project is still in the development stages,
    and in our experience is one one of the more difficult components in this
    kind of systems. In this stage of the evolution, we first need to assign a
    good fitness to those structures that keep their balance and not fall apart.
    There are other variables we take in to account, but they will be explained
    later. In this initial stage, each individual in the population has 100
    points so before the simulation starts each of them are viable candidates to
    be an elite % Elite? - JJ
    member of the population regardless of their relative height or
    complexity.

>>>>

\begin{abstract}
In this paper, we propose an evolutionary algorithm that follows an open-ended
evolution approach to generate levels for the Angry Birds video game. The levels
themselves are composed of a set of birds that the player can throw with a
slingshot, a certain amount of pigs that they must destroy, and a given number
of pieces that conform structures that may or may not protect the pigs from
birds thrown at them. The current goal is the generation of diverse structures
that are playable in the game, having the additional characteristics of being
fun and enjoyable. We propose a multi-layered search, first constructing
composite structures from basic blocks, to then build more complex structures
building from these composites. We follow an open-ended evolution approach in
which the evolution of structures is not guided towards a single objective but
is rather free to evolve and generate novelty or diversity. The fitness function
we use to evaluate the proposed levels considers how complex levels have become
and how different they are from the rest of the population. The experiments
conducted show that an evolutionary approach allows the generation of levels
that are novel and interesting to play.
% Another revision to English would be good.
% Done - Mario
\end{abstract}

@@

\section{State of the art}
\label{SoA}
In this Section, we present PCG works relevant to the level generation for
the Angry Birds video game. Renz et al. \cite{Renz, Renz2015TheAB} organized
an AI competition to develop intelligent agents that could play new  Angry
Birds levels better than the best human players. The competition was then
collocated in other AI conferences. To better train and test agents, more
novel and some times, challenging levels must be generated. Stephenson et
al. proposed a related competition this time with the goal of developing a
level generator for the game. Researchers have proposed several generators
in the following years of competition. M. Katina et al.  \cite{Kaidan2015}
propose a model in which the algorithm controls the generation of content by
using the skill of a player, while she is actively playing the game. The
moment the player finishes a certain level, the algorithm obtains the score
and generates the next level to play. Y. Jiang et al. \cite{Jiang2017}
proposed the use of predefined letter patterns and combined them to create a
small pool of words and letters used to create levels with text on them.
Users would play the game, and after that,   asked them if they liked the
layout of the Levels.  Several researchers follow a search-based approach.
Calle et al. \cite{lauracalle} propose a free-search method to generate
levels.  For this, lists of blocks are placed in the game by the algorithm,
and those maintaining their positions against gravity are selected to build
structures. The resulting structures continue evolving into more complex
levels. That work is the basis of our current proposal, and since we are
using a modified version of the Angry Birds game, that can obtain data of a
level after a simulation.

    \section{Proposed Method}
    \label{prop_method}

In this Section, we describe the approach used in this work, explaining
first, the restrictions of the game, then giving a quick overview of how we
create new composites to be used by the algorithm, then we define how candidate
solutions are to be represented by the algorithm. We then describe our method to
evaluate candidate solutions. Finally, we present details of the evolutionary
algorithm.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[width=80mm]{Images/list_pieces.png}}
\caption{List of blocks in the angry birds game.}
\label{piece_list}
\end{figure}

The Angry Birds level generation track is a competition where the objective is
to generate levels that are \textit{fun} and \textit{enjoyable}; they also have
to be challenging but not impossible. The rules for the competition specify that
a configuration file must be used to generate the levels of the game in a
limited amount of time. The file contains several restrictions;
namely:\begin{itemize}
\item   The number of \textit{different}
levels that must be generated. 
\item The materials that are \textit{not allowed}.

@@

Structures are represented in a way similar to Togelius et al.
\cite{togelius2016Representationsforsearch-basedmethods}. We can model a
structure as a collection of composites, each one selected from a pool of
currently available composites. Initially, the pool contains only the 11
original blocks represented as single-element composites. As new composites are
automatically (or manually) generated from the original blocks, we insert them
to the pool and assign each one an ID. We can model the chromosome of a
structure, as shown in Figure \ref{old_chrom} where each gen is an ID from a
composite in the pool. A chromosome can include the same ID multiple times.
    % I don't understand this too well, maybe rephrasing would clarify - JJ
    % Done - Mario

@@

    These structures represent individuals in the evolutionary algorithm. Once
    an individual is created it is ready to be evaluated in a simulation on the
    game environment. %In Laura's work, we tried to avoid this because it took a long time. Is there any procedure to check constraints and not evaluate them? - JJ
    % I will comment that we could work on that. At the moment each one is evaluated. - M

    Using a similar approach as the composite generator, all
    individuals have a maximum of ten seconds to maintain balance or reach
    absolute stability. If that happens  it means one of two
    things, first the generated structure is completely stable or second, the
    tower fell to either side and the pieces were destroyed or stopped moving by
    gravitational pull, when either of this happens by more than two seconds the
    simulation ends regardless of time remaining and the next individual is
    simulated.

    The evaluation process for this project is still in the development stages,
    and in our experience is one one of the more difficult components in this
    kind of systems. In this stage of the evolution, we first need to assign a
    good fitness to those structures that keep their balance and not fall apart.
    There are other variables we take in to account, but they will be explained
    later. In this initial stage, each individual in the population has 100
    points so before the simulation starts each of them are viable candidates to
    be an Elite % Elite? - JJ
    % But I think the problem remains, there is no difference even if they all have
    % zero points. But randomess in the selection coyuld do that. - Mario
    member of the population regardless of their relative height or
    complexity.
